### 2015-09-12 paper/main.tex
Fix #5 - remove line on CUSFTA
Also add Dix Carneiro (2014) and Costinot/Rodriguez-Clare(2014) to intro

The economic benefits of free trade are arguably one of the most uncontroversial results of economic research, both theoretically and empirically. However, to this date, free trade is by no means uncontroversial in the public sphere, as is evidenced by the fierce opposition that the proposed transatlantic free trade agreement between the US and Europe is facing. Hence, international trade has remained an active field in economic research, a field which has seen major advancements in the past two decades in incorporating firm-level heterogeneity coupled with consumer love of variety into trade models that can account for the firm-level responses to increasing trade openness and the large share of intra-industry trade in the international flow of goods and services \footnote{A comprehensive survey of trade models with love of variety preferences and firm-level heterogeneity can be found in \citet*{MelitzTrefler2012}.}. This new vintage of trade models predicts additional welfare gains from trade stemming from reallocations of production to more productive firms (as in \citealp{Melitz2003}) or increases in firms' efforts to innovate (as in \citealp{GrossmanHelpman1990}). \\
While these new models of international trade are well grounded in empirical evidence coming from micro data, there are surprisingly few tests of the model predictions for aggregate variables which are decisive for the predicted welfare gains from trade. Recently, \citet{Arkolakisetal2012a,Arkolakisetal2012b} call into question the importance of firm-level heterogeneity by showing that in a lass class of trade models, the additional welfare gains are fairly small and actually even smaller if consumers don't have CES utility. The response of \citet{MelitzRedding2013} shows that there is still considerable disagreement over how to theoretically evaluate the additional welfare gains from firm selection, a fact that motivates us to test the model directly in aggregate data on prices, markups and productivity. To do so, we estimate the effects of trade liberalization on the competitive environment in manufacturing markets of the member countries of the North American Free Trade Agreement (NAFTA). We employ an estimation procedure based on the \citet{MelitzOttaviano2008} model introduced by \citet{Chen2009}, which to our knowledge is the only empirical application of a model with firm-level heterogeneity on aggregate data. \citet{Chen2009} derive estimable regression equations from the model's equilibrium conditions that allow us to test the effects of trade openness on relative price levels, markups and labor productivities of two trading partners. It is further possible to differentiate between the effects of trade in the short run, which, in the model, refers to an economy without relocation decisions for firms, and in the long run, when firms are free to choose their home market for production. However, as the underlying model is static, no direct results on the time path of the impact of trade liberalization can be obtained. We try to address this issue by dividing our sample in ways that make it more amenable to a model-based estimation. Contrary to \citet{Chen2009}, we directly observe tariff rates between the three countries in our sample and hence use those as a direct measure of trade openness. Additionally, we test for the effects of third-country trade openness on the relative performance of two countries that are linked through trade, predictions for which can be derived from the multi-country version of the Melitz and Ottaviano model. Our dataset comprises of nine manufacturing sectors in Canada, Mexico and the US, covering the time period from the introduction of NAFTA in 1994 up to 2006, which gives us reason to believe that we are able to capture the long run effects of policy changes even in industries with low firm churning rates\footnote{Given that the US already engaged in a free-trade agreement with Canada in 1988 (CUSFTA), we also estimate the short-run effects of this agreement prior to NAFTA and the results hold}. \\

>>>>

The economic benefits of free trade are arguably one of the most uncontroversial results of economic research, both theoretically and empirically. However, to this date, free trade is by no means uncontroversial in the public sphere, as is evidenced by the fierce opposition that the proposed transatlantic free trade agreement between the US and Europe is facing. Hence, international trade has remained an active field in economic research, a field which has seen major advancements in the past two decades in incorporating firm-level heterogeneity coupled with consumer love of variety into trade models that can account for the firm-level responses to increasing trade openness and the large share of intra-industry trade in the international flow of goods and services \footnote{A comprehensive survey of trade models with love of variety preferences and firm-level heterogeneity can be found in \citet*{MelitzTrefler2012}.}. This new vintage of trade models predicts additional welfare gains from trade stemming from reallocations of production to more productive firms (as in \citealp{Melitz2003}) or increases in firms' efforts to innovate (as in \citealp{GrossmanHelpman1990}). To some extent, these new models of trade also help reconcile the unambiguously positive stance of economic researchers on trade liberalization with the public opposition to it -- models taking into account explicitly the heterogeneity across agents of firms within a country show that while on aggregate there are significant efficiency gains from free trade, there are also firms and workers who will lose out individually, and can only benefit from a trade liberalization if either the aggregate gains are redistributed in some way to ensure a Pareto improving allocation, or if they can benefit from the reallocation of production to more productive firms by switching to those firms. \citet{DixCarneiro2014} builds a structural model of the Brazilian labor market to estimate the labor market effects of trade liberalization and finds that, depending on the assumptions about capital mobility, the reallocation of workers across sectors can take up to 30 years. \\
While these new models of international trade are well grounded in empirical evidence coming from micro data, there are surprisingly few tests of the model predictions for aggregate variables which are decisive for the predicted welfare gains from trade. Recently, \citet{Arkolakisetal2012a,Arkolakisetal2012b} call into question the importance of firm-level heterogeneity by showing that in a lass class of trade models, the additional welfare gains are fairly small and actually even smaller if consumers don't have CES utility. The response of \citet{MelitzRedding2013} shows that there is still considerable disagreement over how to theoretically evaluate the additional welfare gains from firm selection, and \citet{CostinotRodriguez2014} review the effects of trade liberalizations in a wider class of new trade models to highlight the importance of the market structure under consideration -- depending on whether a one- or multi-sector model is used and the degree of competition assumed, gains from trade are estimated to range from 4\% to 40\% of non-free-trade welfare. These facts m motivate us to test the Melitz-Ottaviano model directly in aggregate data on prices, markups and productivity. To do so, we estimate the effects of trade liberalization on the competitive environment in manufacturing markets of the member countries of the North American Free Trade Agreement (NAFTA). We employ an estimation procedure based on the \citet{MelitzOttaviano2008} model introduced by \citet{Chen2009}, which to our knowledge is the only empirical application of a model with firm-level heterogeneity on aggregate data. \citet{Chen2009} derive estimable regression equations from the model's equilibrium conditions that allow us to test the effects of trade openness on relative price levels, markups and labor productivities of two trading partners. It is further possible to differentiate between the effects of trade in the short run, which, in the model, refers to an economy without relocation decisions for firms, and in the long run, when firms are free to choose their home market for production. However, as the underlying model is static, no direct results on the time path of the impact of trade liberalization can be obtained. We try to address this issue by dividing our sample in ways that make it more amenable to a model-based estimation. Contrary to \citet{Chen2009}, we directly observe tariff rates between the three countries in our sample and hence use those as a direct measure of trade openness. Additionally, we test for the effects of third-country trade openness on the relative performance of two countries that are linked through trade, predictions for which can be derived from the multi-country version of the Melitz and Ottaviano model. Our dataset comprises of nine manufacturing sectors in Canada, Mexico and the US, covering the time period from the introduction of NAFTA in 1994 up to 2006, which gives us reason to believe that we are able to capture the long run effects of policy changes even in industries with low firm churning rates. \\
